---
title: "SDS 291 - STAT2 R Companion Chapter 4"
output: html_notebook
---

## Section 4.6: Randomization Test for Predictors (01-30-24)

#### load libraries + data
```{r warning= False}

library(Stat2Data)
library(infer)
library(tidyverse)

data("SATGPA")
head(SATGPA)

```

#### generate the randomization distribution

* `specify()` allows us to specify that we're interested in the relationship between the two variables and the y ~ x. 
* `hypothesize()` sets the null hypothesis is true (using infer means we must make this explicit). 
* `null = "independence` tells it our assumption is the two variables have no relationship. 
* `generate()` creates XX number of reps, and we specify that the repetitions are permutations of the same data. 
* `calculate()` will calculate the slopes of these regression lines (aka the r values)
```{r}

null_dist = SATGPA %>%
  specify(GPA ~ VerbalSAT) %>% 
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")

visualize(null_dist)

```

#### generate p-values

calculate the original data's correlation
```{r}

observed_slope = SATGPA %>%
  specify(GPA ~ VerbalSAT) %>%
  calculate(stat = "slope")

```

generate the p-value from the distribution of the random dist's slopes
```{r echo= FALSE}

get_p_value(x = null_dist, obs_stat = observed_slope, direction = "both")

```

visualize the p-value distributions
```{r}

visualise(null_dist) +
  shade_p_value(obs_stat = observed_slope, direction = 'both')

```

#### randomization test for the correlation

generate a randomized test for the correlation between the two variables. 
```{r echo=FALSE}

## change the calculate value to correlation instead of slope
null_correlation_dist = SATGPA %>%
  specify(GPA ~ VerbalSAT) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "correlation")

observed_correlation = SATGPA %>%
  specify(GPA ~ VerbalSAT) %>%
  calculate(stat = "correlation")

get_p_value(null_correlation_dist, observed_correlation, direction = "both")

```

visualise the correlation values
```{r}

visualise(null_correlation_dist) +
  shade_p_value(obs_stat = observed_correlation, direction = "both")

```

#### inference in multiple regression

* adapt your formula from the `specify()` function to instead follow the formula of the multiple regression model.
* the pipe gets the `fit()` instead of `calculate()`
```{r echo=FALSE}

null_models = SATGPA %>%
  specify(GPA ~ VerbalSAT * MathSAT) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  fit()

observed_model = SATGPA %>%
  specify(GPA ~ VerbalSAT * MathSAT) %>%
  fit()

get_p_value(null_models, obs_stat = observed_model, direction = "both")

```

visualize the model values
```{r}

visualize(null_models) +
  shade_p_value(obs_stat = observed_model, direction = "both")

```

#### controlling randomness

3 permutations of the correlation values --> notice how the first and second versions have different correlation values
```{r echo= FALSE}

SATGPA %>%
  specify(GPA ~ VerbalSAT) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 3, type = "permute") %>%
  calculate(stat = "correlation")

SATGPA %>%
  specify(GPA ~ VerbalSAT) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 3, type = "permute") %>%
  calculate(stat = "correlation")

```

setting the seed will prevent different random values when you run the same permutations
```{r message=FALSE}

set.seed(11)
SATGPA %>%
  specify(GPA ~ VerbalSAT) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 3, type = "permute") %>%
  calculate(stat = "correlation")

set.seed(11)
SATGPA %>%
  specify(GPA ~ VerbalSAT) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 3, type = "permute") %>%
  calculate(stat = "correlation")

```






