---
title: "Kumar: SDS 291 - In-Class Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Class 2 | January 31, 2024

### SET UP 

#### load libraries
```{r}

library(Stat2Data)
library(ggplot2)
library(broom)
library(equatiomatic)

```

#### load data
```{r}

## load data
data("BaseballTimes2017")
head(BaseballTimes2017)

```

#### generate visualization
```{r}

plot = ggplot(BaseballTimes2017, mapping = aes(x = Pitchers, 
                                               y = Time)) + 
  geom_point() +
  theme_bw() + 
  geom_smooth(method = lm, se = FALSE, formula = y ~ x)

plot

```

### EXERCISE 1

a: Fit the regression model visualized above
```{r}

baseball_model = lm(Time ~ Pitchers, BaseballTimes2017)

```
b: Print the model summary
```{r}

summary(baseball_model)

```
c: Interpret the model's slope coefficient in a sentence. 

This model indicates that for each increase in pitchers by one during a game, the duration of the game tends to last 8.017 minutes more. 

### EXERCISE 2
Use the fitted model’s equation to compute the residual error for Game #10, between the Los Angeles Angels and the Seattle Mariners, a game where 9 pitchers were used (feel free to use pen and paper when working with the equation). Finally, check your work using the `augment()` function from the `broom()` package to inspect the precise residual error.

#### estimate on your own

isolate the coefficients
```{r}

baseball_coefs = coef(baseball_model)
baseball_coefs

```

generate the estimated time
```{r}

## Game 10 time = slope * # of pitchers + error
g10_time_est = 8.017241*9 + 124.068966
g10_time_est

```

#### calculate the residual

actual time: 
```{r}

## access baseball times at row 10, column 7 (game 10, time)
g10_time = BaseballTimes2017[10, 7]
g10_time

```

calculate the residual by subtracting the estimate from the actual value: 
```{r}

g10_resid = g10_time - g10_time_est
g10_resid

```


#### check your work

calculate the residuals for all rows from the computer's model
```{r}

residuals = augment(baseball_model)

```

isolate game 10's *actual* residual
```{r echo=FALSE}

## remember that R indexes starting at 1, not 0 (fucked in the head)

## (row 10), residuals column (col 4)
residuals[10,4]

```

## Class 3 | February 2, 2024


