---
title: "SDS 291 - STAT2 R Companion Chapter 10: Contrasts of Coefficients"
output: html_notebook
---

## Section 10.2: A Companion to the Statistical Slueth, Chapter 10.2 (03-05-24)

#### load libraries
```{r message=FALSE}

library(Sleuth2)
library(ggplot2)
library(moderndive)
library(equatiomatic)
library(dplyr)

```

#### load data
```{r}

bat_energy = Sleuth2::case1002
head(bat_energy)

```

#### visualize the data
```{r}

ggplot(data = bat_energy, mapping = aes(x = log(Mass), 
                                        y = log(Energy),
                                        color = Type)) + 
  geom_point(size = 2) + 
  geom_parallel_slopes(se = F) + 
  scale_x_continuous(breaks = log(c(10, 50, 100, 500)),
                     labels = c(10, 50, 100, 500)) +
  scale_y_continuous(breaks = log(c(1, 5, 10, 50)),
                     labels = c(1, 5, 10, 50)) +
  theme(legend.position.inside = c(0.75, 0.2)) +
  theme_bw()

```

#### meet the model
```{r}

bat_model = lm(log(Energy) ~ log(Mass) + Type, data = bat_energy)
summary(bat_model)$coefficients

```

```{r}

extract_eq(bat_model, use_coefs = T)

```

because the baseline group here is non echo locating bats, and we want to determine the difference between echo-locating bats and echo-locating birds, we have to either

* change the model's reference group
* use a contrast of coefficients

#### changing the reference level
```{r}

bat_energy = bat_energy %>%
  mutate(Type = factor(Type, levels = c("echolocating bats", ## this is the first group, so it becomes the baseline
                                        "non-echolocating bats",
                                        "non-echolocating birds")))

head(bat_energy)

```

```{r}

echolocating_baseline = lm(log(Energy) ~ log(Mass) + Type, data = bat_energy)
summary(echolocating_baseline)$coefficients

```
interpreting these coefficients: 

* `(Intercept)`: the average log energy consumption of a non-echolocating bat with body mass 0 log grams
* `log(Mass)`: the average change in log energy consumption for each additional log gram of body mass
* `Typenon-echolocating bats`: the difference in average log energy consumption for non-echolocating bats compared to baseline group echolocating bats. 
* `Typenon-echolocating birds`: the difference in average log energy consumption for echolocating birds compared to baseline group echolocating bats. <-- this is the value we're interested in. 

#### contrasts of coefficients
```{r}

bat_energy = Sleuth2::case1002

```

a contrast is a weighted linear combination of the model's coefficients which sum to 0. our model's 4 existing coefficients are: 

```{r echo=FALSE}

extract_eq(bat_model, use_coefs = T)

```
we want to contrast the 3rd and 4th coefficients (`Typenon-echolocating birds` and `Typeecholocating-bats`). 

the general formula for a contrast is: 
$$g = C_0\hat{\beta}_0 + C_1\hat{\beta}_1 + ... + C_k\hat{\beta}_k$$

so if we give coefficients $\hat{\beta}_0$ and $\hat{\beta}_1$ weights of 0, $\hat{\beta}_3$ a weight of 1, and $\hat{\beta}_4$ a counterweight of $-1$:
$$g = (0 \cdot -1.58) + (0 \cdot 0.81) + (1 \cdot 0.1) + (-1 \cdot 0.08)$$
$$g = 0.01 - 0.08$$
$$g = 0.023$$

to determine if $g$'s value is unusual, we can test it in a hypothesis test. however, this requires access to $g$'s standard error. **the variance of a contrast is the sum of each individual coefficient’s variance, plus negative two times the covariance of the two coefficients**. the square root of this value is $g$'s $SE$. this means we need to generate a co/variance matrix from our model:
```{r}

co_variance_matrix = vcov(bat_model)
co_variance_matrix

```

the variance of each coefficient is on the diagonal, and the covariance between each pair is on the off-diagonals: 
```{r}

var_b3 = co_variance_matrix[3, 3]
var_b4 = co_variance_matrix[4, 4]
cov_b3_b4 = co_variance_matrix[3, 4]

var_g = var_b3 + var_b4 + -2*(cov_b3_b4)
var_g

```

take the square root of this value to find the standard error of $g$: 
```{r}

SE_g = sqrt(var_g)
SE_g

```

construct the test statistic
```{r}

## difference between (NE birds and NE bats) - {E bats and NE bats}
g = coef(bat_model)[3] - coef(bat_model)[4]

t = g / SE_g
t

```

test the hypothesis
```{r}

p_value = 2 * pt(abs(t), df = 20 - 4, lower.tail = F)
p_value

```

